Holomorphic field theories and higher algebra

### 2024-09-19 draft.tex
merge

\DeclareMathOperator{\Schw}{Schw}

@@

\def\brian#1{{\textcolor{blue!75!red}{BW: {#1}}}}
\def\owen#1{{\textcolor{violet!50!black}{OG: {#1}}}}

@@

\owen{mention some history (e.g., Manin, Donaldson-Thomas, Nekrasov) and some recent work (CPTVV, Pridham, Kapranov, etc) but then explain our focus: higher analogs of VOAs and finding mathematical analogs of physics}

@@

The most important field theories in physics involve a metric on a manifold (whether Riemannian or Lorentzian), and hence geometry has long had a clear role in physics.
In recent decades attention has expanded to include {\em topological} field theories (or TFTs), 
in which --- to simplify --- only the underlying smooth topology of the manifold matters.
There are, however, theories that depend on a complex structure on the manifold,
and as one might hope, these are particularly beautiful, 
just as complex variables is a particularly beautiful wing of analysis.
In essence, such {\em holomorphic} field theories have variational partial differential equations (i.e., ``equations of motion'') that are holomorphic.
These theories admit independent motivations from mathematics and physics,
which we sketch before giving more careful definitions.

The mathematical motivation is simple:
for most classical holomorphic field theories, the solutions to the equations of motion form a moduli space of natural interest to complex geometers,
and one might hope that the quantization of this moduli space is equally interesting.
A well-known example is holomorphic Chern-Simons theory,
which lives on a Calabi-Yau 3-fold~$X$ and involves the moduli space $\RR\Bun_G(X)$ of holomorphic principal $G$-bundles on~$X$.
(We add the prefix $\RR$ because it is best to use the {\em derived} moduli space.
We do not assume the reader is versed in derived geometry and will offer motivations, glosses, and references throughout the survey.)

As a more mundane example, let $\RR\Hol(X,Y)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $Y$.
(For $Y = V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
In derived geometry one can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a field theory known as the $\beta\gamma$ system with target~$V$.\footnote{The name is silly: it's due to the original notation people used.}
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of the AKSZ $\sigma$-models (such as the Poisson $\sigma$-model that appears in Kontsevich's construction of deformation quantization); 
the relevant moduli space being the derived mapping space between complex manifolds.

A physical motivation comes from supersymmetric field theories, 
which (loosely speaking) are theories on $\RR^n$ whose symmetries include not only the isometries of $\RR^n$, but an extension to a super Lie group known as a super Poincar\'e group.
%The relevant extension is often labeled by $\cN=1, 2,$ and so on, 
%with larger $\cN$ denoting a bigger supergroup. 
Thanks to the enhanced symmetry, such theories can often be understood in greater detail than non-supersymmetric theories.
The relevance to this survey is that supersymmetric theories admit ``twists'' (in essence, deformations) that are holomorphic field theories, 
even when they do not admit twists to TFTs \cite{CosHol}. 
(The famed A- and B-models of mirror symmetry are TFTs produced by twisting supersymmetric thories.)
As an example, four-dimensional minimally supersymmetric Yang-Mills theory admits no topological twist, 
but it does admit a twist to holomorphic field theory whose moduli space is described, in part, by the moduli space of holomorphic $G$-bundles~$\RR \Bun_G(X)$.

It is natural to wonder how the myriad tools and computations deployed on a supersymmetric theory carry over to this twist 
and, conversely, whether information can flow back from a holomorphically twisted theory to the original supersymmetric theory, 
much as TFTs have enriched the understanding of supersymmetric theories.
In short, a holomorphic twist contains the ``holomorphic sector'' of the untwisted theory just as a topological twist contains its ``topological sector.''
To aid translation between mathematics and physics, 
it may be helpful to know that the role of derived geometry on the mathematical side is mirrored by the use of the BRST/BV formalism on the physics side;
they are closely related and serve similar purposes.

As a final motivational remark, we note in one complex dimension (i.e., working on Riemann surfaces), 
holomorphic field theories appear as chiral conformal field theories (CFTs) on Riemann surfaces. 
The study of CFTs--- whose manifestations in mathematics include vertex algebras, affine Lie algebras, and loop groups --- have had a strong impact in mathematics,
so it is natural to explore higher-dimensional analogs.
So far, most mathematical effort in this direction has focused on (essentially) classical aspects of gauge theory,
but the quantum aspects are profoundly interesting.
We expect that these holomorphic examples will be particularly tractable and beautiful but also useful in offering insights even for non-holomorphic theories.

\subsection{The aims of this survey}

Over the last few decades there has been a vigorous dialogue between complex geometry and theoretical physics,
around topics like mirror symmetry, enumerative algebraic geometry, moduli of Calabi-Yau manifolds, the geometric Langlands program, and integrable systems, among others.
Covering all these topics would be difficult.
Our survey's focus will be narrow:
it is upon higher dimensional versions of the relationship between geometry and algebraic structures like Lie algebras, Poisson algebras, and vertex algebras.
We will emphasize analogies to TFTs (such as the Gerstenhaber algebras appearing in mirror symmetry) and CFTs (such as the affine Lie algebras relating to moduli of $G$-bundles on Riemann surfaces),
but our primary interest is on complex manifolds above dimension one.
Experience suggests that higher dimensions require higher algebra,
i.e., algebraic methods that involve cochain complexes.
Thus differential graded (dg) algebras and categories,
as well as fellow travelers like derived geometry, appear.

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK}, \cite{Keviny things}, \cite{GaiottoEtAl}, \cite{Paquette}, \cite{Vicedo}, \cite{BeemBZEtAl}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.

The material below focuses on connections with this work,
and the examples from physics have a perturbative flavor.
We will not discuss important directions like mirror symmetry, the geometric Langlands program, the BPS/CFT correspondence, Donaldson-Thomas theory, or other areas that are more global or more directly geometric,
although we expect there are rich connections with the topics we do discuss.
Not including these topics is primarily because our core theme already swelled to more pages than we expected, 
because a concise discussion of any of those topics does not yet seem possible with respect to our theme,
and because of our lack of background in these areas.
We hope the ideas surveyed here do connect profitably to these topics.
\owen{Is that an overly negative paragraph?}

\subsection{How to read this paper}

Section~\ref{sec: HFT} is devoted to describing holomorphic field theories, primarily by offering a bounty of examples, including holomorphic analogs of gauge and gravity theories.
We outline as well how to obtain holomorphic field theories from supersymmetric field theories,
a rich source of examples to explore.

The next three sections are, in a sense, the core of the survey,
as they describe increasingly general types of higher algebras and how these capture the observables of holomorphic field theories.
Section~\ref{sec: algebra} revisits the relationship between quantum mechanics and associative algebras,
chiral conformal field theory and vertex algebras,
and holomorphic field theories and higher-dimensional generalizations of vertex algebras.
Section~\ref{sec: config} offers a useful framework for studying these higher vertex algebras,
using operads,
and it develops powerful analogies to the little $n$-disks operad $E_n$ that plays an important role in contemporary topology.
Section~\ref{sec: fact alg} introduces factorization algebras,
providing a local-to-global principle to describe the observables on complex manifolds beyond affine spaces.
In particular, by pushing forward factorization algebras,
one can relate field theories of different dimension,
i.e., a version of the physicist's ``compactification.''

A reader could read this core portion independently of the rest of the paper,
as it focuses on algebra and geometry and requires no substantial knowledge of field theory {\it per se}.

Section~\ref{sec: renorm} describes recent work that allows one to produce examples:
there is now a convenient framework for renormalization of holomorphic field theories on $\CC^n$,
and so the field theories of Section~\ref{sec: HFT} generate higher algebras of the form described in Sections~\ref{sec: algebra},~\ref{sec: config}, and~\ref{sec: fact alg}.
The reader could skip to this section directly from Section~\ref{sec: HFT}.

We conclude the paper by describing the kind of fascinating mathematical questions
that arise by taking physical conjectures (e.g., dualities of supersymmetric theories) and twisting them into conjectures about holomorphic field theories. 
Section~\ref{sec: seiberg} explores the holomorphic cousin of Seiberg duality,
which suggests a remarkable correspondence between a gauge theory with group $SL(N)$ and a gauge theory with group $SL(M)$, where $M \neq N$.
It is a kind of enhancement of Weyl's invariant theory and has close connections with topics like homological projective duality.

\subsection{Acknowledgements}

\owen{To thank: Kevin Costello, Si Li, Davide Gaiotto, Ed Segal, Chris Elliott, Pavel Safronov, Kapranov, Hennion, Matt Szczesny, Richard Eager, Ingmar Saberi, Johannes Walcher, Jie Zhou, Minghao Wang, Nik Rozenblyum}

\section{Holomorphic field theories}
\label{sec: HFT}

\owen{I rearranged this section somewhat. Our discussion of examples did not involve any BRST/BV stuff, so I put it right after the impressionistic opening. The "definitions" section seemed to be purely perturbative and to use notions from derived geometry, so I put it next. It should be fleshed out a little (e.g., a mention of the "fundamental theorem of deformaiton theory" and maybe show how to write one of the examples in that language).}

We begin by sketching what a classical field theory is, somewhat impressionistically, and then offer some concrete examples.

In physics, a classical field theory is usually described in terms of a few pieces of data:
\begin{itemize}
\item a manifold $M$ that plays the role of ``spacetime,''
\item a space of fields, such as sections of some fiber bundle $F \to M$ or connections on a principal bundle on $M$, and
\item a system of PDE for the fields, known as the equations of motion.
\end{itemize}
These equations of motion are variational: 
they are the Euler-Lagrange equations of an action functional on the space a fields.
An action functional is written as an integral over $M$ of some polydifferential operator on the fields that takes values in densities of~$M$.
A few examples will illuminate this sketch.

\begin{eg}[Massless scalar field theory]
\label{eg: massless scalar}
Let $M$ be a smooth manifold and equip it with a Riemannian metric~$g$.
Let $\Delta_g$ denote the associated Laplace-Beltrami operator, 
and let $\d{\rm vol}_g$ denote the associated volume form.
Let the space of fields be the smooth functions $C^\infty(M)$.
A smooth function $\phi$ is called, in physics, a {\em scalar} field.
The action functional
\[
S(\phi) = \int_M \phi \Delta_g \phi \, \d{\rm vol}_g
\]
corresponds to the equation of motion $\Delta_g \phi = 0$. 
Thus this scalar field theory has the moduli space of {\em harmonic functions} as its space of solutions.
\end{eg}

The preceding example exhibits how a moduli space of central interest in differential geometry --- the harmonic functions --- appears naturally as a basic example in physics (the ``free massless scalar field'').
More complicated moduli spaces show up too.

\begin{eg}[$\phi^4$ theory]
As a small variant of the example just given,
consider the action functional
\[
S(\phi) = \int_M \phi \Delta_g \phi + \frac{1}{4!} \phi^4\, \d{\rm vol}_g.
\]
Its equation of motion is the nonlinear PDE 
\[
\Delta_g \phi + \frac{1}{3!} \phi^3 = 0.
\] 
Thus this scalar field theory has a more complicated, nonlinear moduli space of solutions.
\end{eg}

We now describe an example of a theory whose space of solutions is a moduli space of interest in geometric topology.

\begin{eg}[Chern-Simons theory]
Pick a compact Lie group $G$ and equip it with an invariant metric,
which induces a $G$-invariant inner product $\kappa$ on its Lie algebra~$\fg$.
(The standard choice is the Killing form.)
Take an oriented smooth 3-manifold $M$, 
and consider the space of connections on the trivial principal $G$-bundle on $M$.
This space of fields is isomorphic to $\Omega^1(M, \fg)$, the $\fg$-valued one-forms, 
by viewing $\d + A$ as a connection where $A \in \Omega^1(M, \fg)$.
The equation of motion is the zero-curvature equation
\[
F_A = 0,
\]
which can be expressed equivalently as the Maurer-Cartan equation
\[
\d A + \frac{1}{2} [A,A] = 0.
\]
These equations arise as the Euler-Lagrange equation for the action functional
\beqn
\label{eqn:CS}
CS(A) = \frac12 \kappa(A, \d A) + \frac16 \kappa(A, [A,A]),
\eeqn
the integral of a Chern-Simons form.
Hence this theory is known as {\em Chern-Simons theory}.\footnote{Note that $M$ needs to be an oriented 3-manifold for this action functional to make sense.
The equation of motion exists for a principal bundle on any manifold of any dimension,
but it is only variational for oriented 3-manifolds.}
These equations do not depend on a metric on $M$ or any other geometric structure,
and hence define a ``topological'' field theory in the sense that they are independent of geometry.
This theory is interesting because the moduli space of solutions consists of the flat $G$-bundles on $M$,
a topic of deep interest in geometry and topology.
For physicists, it is interesting because it does not involve geometry, 
by contrast to earlier examples of field theories,
and yet exhibits rich phenomena (notably applications to the quantum Hall effect and 3-dimensional gravity).
\end{eg}

\begin{rmk}
In the Chern-Simons example, there is an important aspect we did not yet acknowledge:
we need to identify different solutions related by gauge transformations.
In other words, the correct moduli space is the quotient stack
\[
\left[ \{ A \,:\, F_A = 0\}/\Map(M,G) \right].
\]
The usual mode of description of a field theory, by specifying fields and an action functional (or equations of motion), 
must be supplemented by specifying whether to quotient by some group of symmetries of the space of solutions.
More generally, a field theory ought to present some kind of derived moduli stack.
\end{rmk}

We are now in a good position to offer an example of a {\em holomorphic field theory}:
the holomorphic analog of Chern-Simons theory.

%In a {\em holomorphic theory} one starts with a complex manifold $X$ as the spacetime.
%%This means, in part, that the complexified tangent bundle of $X$ decomposes as a sum of vector bundles $\T_X \oplus \Bar{\T}_X$ where $\T_X$ (respectively $\Bar{T}_X$) is the holomorphic (respectively anti-holomorphic) tangent bundle.
%%In local coordinates sections of $\T_X$ have the form $f^i \del / \del z_i$ whereas sections of $\Bar{\T}_X$ have the form $f^i \del / \del \zbar_i$. 
%Globally, holomorphic theories are described like in the usual case by mapping spaces or spaces of connections.
%The key feature in a holomorphic theory is that the equations of motion dictate that fields be holomorphic in an appropriate sense.
%Perturbatively, where fields are sections of some (possibly graded) holomorphic vector bundle $V$, we can make this precise. 
%Prescribing a holomorphic structure on a vector bundle $V$ is equivalent to the data of a $\dbar$-operator
%\beqn
%\dbar \colon \Omega^{0,q}(X, V) \to \Omega^{0,q+1}(X,V) 
%\eeqn
%where $\Omega^{0,q}(X,V)$ is the space of smooth sections of the vector bundle $\wedge^q\Bar{T}_X^* \otimes V$. 
%Holomorphicity of $V$ boils down to the requirement that $\dbar \circ \dbar = 0$.
%The Dolbeault complex of $V$ is
%\beqn
%\left(\Omega^{0,\bu}(X, V), \dbar \right) = \left(\oplus_{q \geq 0} \Omega^{0,q}(X, V)[-q], \dbar \right)
%\eeqn
%where (in the case that $V$ sits in degree zero) the space $\Omega^{0,q}(X,V)$ sits in degree $+q$.
%A standard result in complex geometry is that the Dolbeault complex is a resolution for the sheaf of holomorphic sections of the holomorphic vector bundle $V$.
%%In the case that $V$ is the trivial bundle $\ul{\CC}$ the Dolbeault complex $\Omega^{0,\bu}(X)$ is a resolution for the sheaf of holomorphic functions on $X$.
%
%The fields of a holomorphic theory are given by elements of the Dolbeault complex $\Omega^{0,q}(X,V)$ for some $q$.
%The equations of motion in a holomorphic theory take the form
%\beqn
%\dbar \Phi + \sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
%\eeqn
%where $D_k$ are $k$-ary {\em holomorphic} polydifferential operators.
%This means that in local coordinates $D_k$ only depend on the holomorphic coordinates $\{z_i\}$ and involve only holomorphic derivatives $\del / \del z_i$.

\begin{eg}[Holomorphic Chern-Simons theory]
\label{eg: holCS}
\owen{This feels rather convoluted. Can we make it simpler?}
Pick a {\em complex} Lie group $G$, 
and equip its Lie algebra $\fg$ with a complex-linear, $G$-invariant nondegenerate symmetric bilinear form~$\kappa$.
Let $M$ be a three-dimensional \textit{complex} manifold
equipped with a holomorphic volume form $\Omega \in \Omega^{3,0}(M)$; 
in other words, $M$ is a Calabi--Yau manifold.
(This condition is analogous to being an oriented 3-manifold.)
Given a smooth principal $G$-bundle $P \to M$,
there is a moduli space of holomorphic structures on~$P$.
Suppose there exists at least one holomorphic structure $\cP \to M$,
which has an associated Dolbeault operator (or $\dbar$-connection) $\dbar_{\cP}$.
Any other holomorphic bundle structure determines a Dolbeault operator,
but our reference point $\dbar_{\cP}$ provides an isomorphism
\[
{\rm Conn}^\dbar_G(P) \cong \Omega^{0,1}(M, {\rm ad}(P)),
\]
i.e., we express another $\dbar$-connection as $\dbar_{\cP} + A$ where $A$ is a $(0,1)$-form valued in the adjoint bundle.
The equation of motion is again a Maurer--Cartan equation
\beqn
\dbar_{\cP} A + \frac12 [A,A] = 0 ,
\eeqn
or equivalently the $(0,2)$-term of the curvature vanishes.
This PDE is the Euler--Lagrange equation for the action
\beqn
S_{hCS} (A) = \int_X \Omega \wedge CS(A) 
\eeqn
where $CS(A)$ is exactly as in~\eqref{eqn:CS}.\footnote{Notice that due to the presence of the holomorphic volume form $\Omega$, only the $\dbar A$ component of $\d A$ contributes to the integrand of the action functional, because $\Omega \wedge \partial A$ vanishes.} 
As in the case of Chern-Simons theory, one must quotient by the appropriate group of gauge transformations to obtain the correct moduli space, 
which is here the moduli stack of holomorphic principal $G$-bundles whose underlying smooth structure is given by $P \to M$.
\end{eg}

For a systematic introduction to classical field theory, 
in a style amenable to a complex geometer,
we suggest \cite{FreedDeligne}.
There are many other sources, such as \cite{}, because classical field theory is an old and deep subject.
Recently there is an effort to revisit field theory with insights from derived geometry.
See \cite{AlfonsoYoung?} for motivation and overview, 
but for a detailed development, see \cite{Steffens}.
In Section~\ref{sec: pert dfn} we offer a sharp definition well-suited to perturbative field theory 
(i.e., studying formal deformations of a fixed solution to the equations of motion),
which we will use to produce examples of higher algebras.

@@

In this section we describe several examples of holomorphic field theories,
in the style we used for holomorphic Chern-Simons theory.
Our aim here is to give a sense of how take physical ideas and find holomorphic analogs.

In traditional physics there are a few basic types of field theories, 
and they can be assembled into combinations,
of which the Standard Model offers a nice example.
The basic types are\footnote{For the experts, we remark that we focus above on the field content that motivates or characterizes, and mostly ignore (just for the moment!) issues like gauge transformations and the associated ghost fields.}
\begin{itemize}
\item {\em Scalar} field theories where the fields are maps from $M$ to a vector space $V$, or natural generalizations like taking sections of a vector bundle $V \to M$. The Higgs boson appears as part of a scalar field theory.
\item {\em Fermionic} field theories where the fields are maps from $M$ to an {\em odd} vector space (in the sense of supermathematics), typically a section of a spinor bundle. Such theories are often used to describe matter, such as electron fields.
\item $\sigma$-{\em models} where the fields are maps from $M$ into a manifold~$X$. (One can view these as nonlinear generalizations of scalar field theories.)
\item {\em Gauge} field theories where the fields are connections on a principal $G$-bundle over $M$, with $G$ a Lie group. The Yang--Mills theories for the electroweak and strong forces are examples.
\item {\em Gravity} theories where the fields are metrics on $M$, or something that similarly controls the geometry of $M$. Einstein's theory is a paradigmatic example.
\end{itemize}
The Standard Model involves the first three types,
and the different fields interact in intricate ways.\footnote{For instance, the scalar and fermionic fields are sections of associated bundles to the principal bundle appearing in the gauge theory.}

>>>>

@@

\def\brian#1{{\textcolor{blue!65!red}{BW: {#1}}}}
\def\owen#1{{\textcolor{violet!65!black}{OG: {#1}}}}

@@

\owen{mention some history (e.g., Manin! Donaldson-Thomas, Nekrasov) and some recent work (CPTVV, Pridham, etc) but then explain our focus: higher analogs of VOAs and finding mathematical analogs of physics}

@@

The most important field theories in physics involve a metric on a manifold (whether Riemannian or Lorentzian), and hence geometry has long had a clear role in physics.
In recent decades attention has expanded to include {\em topological} field theories (or TFTs), 
in which --- to simplify --- only the underlying smooth topology of the manifold matters.
There are, however, theories that depend on a complex structure on the manifold,
and as one might hope, these are particularly beautiful, 
just as complex variables is a particularly beautiful wing of analysis.
In essence, such {\em holomorphic} field theories have variational partial differential equations (i.e., ``equations of motion'') that are holomorphic:
the PDE only involves derivatives in the holomorphic coordinates (i.e., the $\partial/\partial z_j$) and has holomorphic coefficients.
These theories admit independent motivations from mathematics and physics,
which we sketch before giving more careful definitions.

The mathematical motivation is simple:
for most classical holomorphic field theories, the solutions to the equations of motion form a moduli space of natural interest to complex geometers,
and it is natural to hope that its quantization is equally interesting.
A well-known example is holomorphic Chern-Simons theory,
which lives on a Calabi-Yau 3-fold~$X$ and involves the moduli space $\RR\Bun_G(X)$ of holomorphic principal $G$-bundles on~$X$.
(We add the prefix $\RR$ because it is best to use the {\em derived} moduli space.
We do not assume the reader is versed in derived geometry and will offer motivations, glosses, and references throughout the survey.)
As a more mundane example, let $\RR\Hol(X,Y)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $Y$.
(For $Y = V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
In derived geoemtry one can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a theory known as the $\beta\gamma$ system with target $V$.
(The name is silly: it's due to the original notation people used.)
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of the AKSZ $\sigma$-models (such as the Poisson $\sigma$-model that appears in Kontsevich's construction of deformatino quantization); 
the relevant moduli space being the derived mapping space between complex manifolds.

A physical motivation comes from supersymmetric field theories, 
which (loosely speaking) are theories on $\RR^n$ whose symmetries include not only the isometries of $\RR^n$, but an extension to a super Lie group known as a super Poincar\'e group.
%The relevant extension is often labeled by $\cN=1, 2,$ and so on, 
%with larger $\cN$ denoting a bigger supergroup. 
Thanks to the enhanced symmetry, such theories can often be understood in greater detail than non-supersymmetric theories.
The relevance to the models we study is that supersymmetric theories admit ``twists'' (in essence, deformations) that are holomorphic field theories, 
even when they do not admit twists to TFTs \cite{CosHol}. 
(The famed A- and B-models of mirror symmetry are TFTs produced by twisting supersymmetric thories.)
As an example, four-dimensional minimally supersymmetric Yang-Mills theory admits no topological twist, but does admit a twist to holomorphic field theory whose moduli space is described, in part, by the moduli space of holomorphic $G$-bundles~$\RR \Bun_G(X)$. 
It is natural to wonder how the myriad tools and computations deployed on the supersymmetric theory carry over to this twist and conversely whether information can flow back from a holomorphically twisted theory to the untwisted one, 
much as TFTs have enriched the understanding of supersymmetric theories.
In short, a holomorphic twist contains the ``holomorphic sector'' of the untwisted theory just as a topological twist contains its ``topological sector.''
It may be helpful to know that the role of derived geometry on the mathematical side is mirrored by the use of the BRST/BV formalism on the physics side;
they are closely related and serve similar purposes.

As a final motivational remark, we note in one complex dimension (i.e., working on Riemann surfaces), holomorphic field theories appear as chiral conformal field theories (CFTs) on two-dimensional manifolds --- whose manifestations in mathematics include vertex algebras, affine Lie algebras, and loop groups --- and have had a strong impact in mathematics.
It is natural to explore higher-dimensional analogs.
So far, most mathematical effort in this direction has focused on classical aspects of gauge theory,
but the quantum aspects are profoundly interesting.
We expect that these holomorphic examples will be particularly tractable and beautiful but also useful in offering insights even for non-holomorphic theories.

The overarching goal of this paper is to introduce the reader to this emerging domain where physics and mathematics meet,
with an emphasis upon how key objects from physics, notably supersymmetric theories and conformal field theories, 
interface with complex geometry in general and with mathematical methods of rather recent vintage,
notably derived geometry and higher algebras.
A community is forming, involving both mathematicians and physicists,
to bring into sharp mathematical form some of the deep insights into gauge theory offered by physics.
There is a lot to explore here.

\subsection{Preliminary notions}

\brian{complex geometry, dolbeault, dg stuff.}

A physical field theory is usually described in terms of a space of fields that forms some possibly nonlinear space, like a mapping space $\Map(M,X)$ or the space of connections ${\rm Conn}_G(M)$ on a manifold. 
There is a manifold, called spacetime in physics, where the fields are some kind of functions out of it;
in the examples just mentioned, it is denoted $M$.
In perturbative field theory, which describes the formal neighborhood of the theory at a point in such spaces, 
a field $\Phi$ is typically described by a smooth section of a vector bundle on spacetime.
The equations of motion that dictate the dynamics of the particular model are encoded by a series of partial differential equations of the form 
\[
\sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\]
where $D_k$ is a $k$-ary polydifferential operator.
As an example, the scalar $\phi^4$-theory on a Riemannian $M$ is described by the equation
\[
\Delta \phi + \frac{1}{3!} \phi^3 = 0 
\]
where $\Phi = \phi$ is simply a smooth function on $M$.
These equations of motion are natural to the Riemannian geometry of the manifold $M$, 
and they can be derived from the Euler--Lagrange equations of the  action functional $\int_M \phi \Delta \phi \, \op{dvol}$.

As another example, consider perturbative Chern--Simons theory whose fields are a principal $G$-bundle $P$ on a smooth three-manifold $M$ equipped with a connection.
The equations of motion is the condition that the connection is flat 
\beqn
F_\nabla = 0 .
\eeqn
Perturbatively, near the trivial flat connection, the fields are simply $\lie{g} = \op{Lie}(G)$-valued one-forms $A \in \Omega^1(M, \lie{g})$.
The condition that perturbed connection $\d + A$ is flat is the Maurer--Cartan equation
\beqn
\d A + \frac12 [A,A] = 0.
\eeqn
These equations of motion are topological in the sense that they only depend on the smooth manifold~$M$.
When $\lie{g}$ is equipped with a non-degenerate, invariant, bilinear form $\kappa$ (for example, if $G$ is semi-simple there is a standard choice) then these equations of motion can be derived from the famous Chern--Simons action functional $\int_M CS(A)$ where
\beqn\label{eqn:CS}
CS(A) = \frac12 \kappa(A, \d A) + \frac16 \kappa(A, [A,A]) .
\eeqn

In a {\em holomorphic theory} one starts with a complex manifold $X$ as the spacetime.
%This means, in part, that the complexified tangent bundle of $X$ decomposes as a sum of vector bundles $\T_X \oplus \Bar{\T}_X$ where $\T_X$ (respectively $\Bar{T}_X$) is the holomorphic (respectively anti-holomorphic) tangent bundle.
%In local coordinates sections of $\T_X$ have the form $f^i \del / \del z_i$ whereas sections of $\Bar{\T}_X$ have the form $f^i \del / \del \zbar_i$. 
Globally, holomorphic theories are described like in the usual case by mapping spaces or spaces of connections.
The key feature in a holomorphic theory is that the equations of motion dictate that fields be holomorphic in an appropriate sense.
Perturbatively, where fields are sections of some (possibly graded) holomorphic vector bundle $V$, we can make this precise. 
Prescribing a holomorphic structure on a vector bundle $V$ is equivalent to the data of a $\dbar$-operator
\beqn
\dbar \colon \Omega^{0,q}(X, V) \to \Omega^{0,q+1}(X,V) 
\eeqn
where $\Omega^{0,q}(X,V)$ is the space of smooth sections of the vector bundle $\wedge^q\Bar{T}_X^* \otimes V$. 
Holomorphicity of $V$ boils down to the requirement that $\dbar \circ \dbar = 0$.
The Dolbeault complex of $V$ is
\beqn
\left(\Omega^{0,\bu}(X, V), \dbar \right) = \left(\oplus_{q \geq 0} \Omega^{0,q}(X, V)[-q], \dbar \right)
\eeqn
where (in the case that $V$ sits in degree zero) the space $\Omega^{0,q}(X,V)$ sits in degree $+q$.
A standard result in complex geometry is that the Dolbeault complex is a resolution for the sheaf of holomorphic sections of the holomorphic vector bundle $V$.
%In the case that $V$ is the trivial bundle $\ul{\CC}$ the Dolbeault complex $\Omega^{0,\bu}(X)$ is a resolution for the sheaf of holomorphic functions on $X$.

The fields of a holomorphic theory are given by elements of the Dolbeault complex $\Omega^{0,q}(X,V)$ for some $q$.
The equations of motion in a holomorphic theory take the form
\beqn
\dbar \Phi + \sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\eeqn
where $D_k$ are $k$-ary {\em holomorphic} polydifferential operators.
This means that in local coordinates $D_k$ only depend on the holomorphic coordinates $\{z_i\}$ and involve only holomorphic derivatives $\del / \del z_i$.

For a simple example, consider the following holomorphic analog of Chern--Simons theory.
The fields are principal $G$-bundles on a three-dimensional \textit{complex} manifold $X$, and the equations of motion give the condition that the principal $G$-bundle is holomorphic.
Perturbatively, near the trivial $G$-bundle, the fields of this holomorphic theory are $\lie{g}$-valued $(0,1)$-forms $A \in \Omega^{0,1}(X, \lie{g})$.
The condition that $\d + A$ prescribes a holomorphic structure is again a Maurer--Cartan equation
\beqn
\dbar A + \frac12 [A,A] = 0 .
\eeqn
Like ordinary Chern--Simons theory, in order for this equation to arise from a principal of least action one needs to fix a non-degenerate, invariant, bilinear form $\kappa$ on $\lie{g}$.
Additionally, one needs to fix a holomorphic volume form $\Omega$ on $X$; in other words, $X$ is a Calabi--Yau manifold.
Then, the equations of motion is precisely the Euler--Lagrange equation for the action
\beqn
S_{hCS} (A) = \int_X \Omega \wedge CS(A) 
\eeqn
where $CS(A)$ is exactly as in \eqref{eqn:CS}.\footnote{Notice that due to the presence of the holomorphic volume form $\Omega$ only the $\dbar$ operator contributes to the kinetic term and so the action can be written as $\int_X \Omega \wedge \left(\frac12 \kappa( A ,\dbar A ) + \frac16 \kappa(A , [A,A]) \right)$.}

\subsection{Precise definitions}

We turn to a formal definition underpinning the formal moduli spaces of the holomorphic field theories that we will be considering.

\begin{dfn}
A {\em holomorphic dg Lie algebra} on a complex manifold $X$ is
\begin{itemize}
\item a graded holomorphic vector bundle $V$ on $X$ whose sheaf of holomorphic functions is $\cV^{hol}$, and
\item a holomorphic differential operator $Q^{hol} \colon \cV^{hol} \to \cV^{hol}$ of cohomological degree one and a holomorphic bidifferential operator $[\cdot,\cdot] \colon \cV^{hol} \times \cV^{hol} \to \cV^{hol}$. 
\end{itemize}
This data is required to endow $(\cV^{hol}, Q^{hol}, [\cdot,\cdot])$ with the structure of a dg Lie algebra.
\end{dfn}

Any holomorphic vector bundle has an associated Dolbeault complex $\Omega^{0,\bu}(X,V)$. 
If $V$ is a holomorphic dg Lie algebra, 
then $\Omega^{0,\bu}(X, V)$ is a dg Lie algebra. 
The differential is of the form $\dbar + Q^{hol}$ and the bracket is given by extending the bracket on $\cV^{hol}$ with the wedge product of differential forms. 
Notice that this dg Lie algebra has the appealing feature that it is the $C^\infty$-sections of a smooth vector bundle on $X$.

Formal moduli spaces arise from the space of Maurer--Cartan elements of a dg Lie algebra.
In our case, this dg Lie algebra will always be of the form $\Omega^{0,\bu}(X, V)$.
For example, the Maurer--Cartan elements of the dg Lie algebra $\Omega^{0,\bu}(X, \fg)$ recovers formal deformations of the trivial holomorphic $G$-bundle.  
This arises from the holomorphic Lie algebra on $X$ whose underlying vector bundle is trivial with fiber $\fg$.

As another example, consider the holomorphic tangent bundle $V = \T_X$. 
The Lie bracket of vector fields endows this with the structure of a holomorphic Lie algebra (with zero differential). 
A Maurer--Cartan element in $\Omega^{0,\bu}(X, \T_X)$ is an element $\mu \in \Omega^{0,1}(X, \T_X)$ satisfying the equation
\[
\dbar \mu + \frac12 [\mu,\mu] = 0 .
\]
This recovers the classical picture describing formal deformations of complex structures on~$X$.

@@

In traditional physics there are a few basic types of field theories, 
and they can be assembled into combinations,
of which the Standard Model offers a nice example.
The basic types are\footnote{For the experts, we remark that we focus above on the core field content, and ignore (just for the moment!) issues like gauge transformations and the associated ghost fields.}
\begin{itemize}
\item {\em Scalar} field theories where the fields are maps from $M$ to a vector space $V$, or natural generalizations like taking sections of a vector bundle $V \to M$. The Higgs boson appears as part of a scalar field theory.
\item {\em Fermionic} field theories where the fields are maps from $M$ to an {\em odd} vector space (in the sense of supermathematics), typically a section of a spinor bundle. Such theories are often used to describe matter, such as electron fields.
\item $\sigma$-{\em models} where the fields are maps from $M$ into a manifold~$X$. (One can view these as generalizations of scalar field theories.)
\item {\em Gauge} field theories where the fields are connections on a principal $G$-bundle over $M$, with $G$ a Lie group. The Yang--Mills theories for the electroweak and strong forces are examples.
\item {\em Gravity} theories where the fields are metrics on $M$, or something that similarly controls the geometry of $M$. Einstein's theory is a paradigmatic example.
\end{itemize}
The Standard Model involves the first three types,
and the different fields interact in intricate ways.\footnote{For instance, the scalar and fermionic fields are sections for associated bundles of the principal bundle appearing in the gauge theory.}

There are holomorphic versions of all these basic types, 
and we now spell out some simple examples.
To give the readers --- possibly a mixed audience of mathematicians and physicists --- some feel for the examples, 
we describe them in several ways: 
in terms of PDE and action functionals but also as moduli spaces.
We do not offer a thorough justification of the dictionary, for which see \owen{cite who?}
